# Getting Started and Getting Help

## 1. Downloading and installing R

## 2. Starting R

## 3. Entering Commands

if expressoin is incomplete, R will prompt for more

## 4. Exiting from R

## 5. Interrupting R

Windows or OS X: Either press the ESC key or click on the stop-sign icon

Linux or Unix: Press Ctrl-C.

## 6. Viewing the Supplied Documentation

```{r}
help.start()
```

## 7. Getting Help on a Function

Use `help` to display the documentation for the function

```{r}
help(functionname)
# Or use the following command
?functionname
```

Use `args` for a quick reminder of the function arguments

```{r}
args(functionname)
```

Use `example` to see examples of using the function

```{r}
example(functionname)
```

## 8. Searching the Supplied Documentation

Use `help.search` to search the R documentation on your computer

```{r}
help.search("pattern")
# Or use the following command
??keyword
```

after search, it will return a list of packages like:

```{r}
Help files with alias or concept or title matching 'adf.test' using regular expression matching:

tseries::adf.test       Augmented Dickey-Fuller Test

Type '?PKG::FOO' to inspect entry 'PKG::FOO TITLE'
```

then use help function 

```{r}
help(adf.test, package="tseries")
```

## 9. Getting Help on a Package

```{r}
help(package="packagename")
```

List all vignettes(additional documents such as introduction, tutorial or reference cards)

```{r}
vignette()
```

See a vignettes for a particular package

```{r}
vignette(package="packagename")
```

Each vignette has a name, use name to view:

```{r}
vignette("vignettename")
```

## 10. Searching the Web for Help

```{r}
RSiteSearch("key phrase")
```

## 11. Find Relevant Functions and Packages

Use following methods:

1. Visit <http://cran.r-project.org/web/views>
2. Visit <https://crantastic.org/>
3. Visit <https://rseek.org/>

## 12. Searching the Mailing Lists

1. Visit <https://rseek.org/>
2. Use

```{r}
RSiteSearch("key phrase")`
```

## 13. Submitting Questions to the Mailing Lists

mail question to r-help@r-project.org
