# Some Basics

## 1. Printing Something

Simply enter the variable name or expression at the command prompt

Use `print` function

`print` doesn't support multiple object, print them one by one

or use `cat`, but `cat` doesn't support complex object 

## 2. Setting Variables

```{r}
x<-3
y<-4
z<-sqrt(x^2+y^2)
```

There is no need for declare or explicitly create variable in R. R is a **dynamically typed language** which means that we can change the variable's data type at will.

global virable assignment

```{r}
x<<-4
```

## 3. Listing Variables

Use `ls` function, use `ls.str` for more details about each variable.

Ordinarily, ls does not return any name that begins with a dot (.). such names are considered hidden. Force `ls` to list everything by setting the `all.names` argument to `TRUE`

```{r}
ls(all.names=TRUE)
```


## 4. Deleting Variables

Use `rm` function

```{r}
x <- 2*pi
rm(x)
```

Remove entire workspace by

```{r}
rm(list=ls())
```

## 5. Creating a Vector

Use the `c(...)` operator to construct a vector from given valeus

a vector can contain either numbers, strings, or logical values but not a mixture

```{r}
c(1,1,2,3,5,8,13,21)
c(1*pi, 2*pi, 3*pi, 4*pi)
c("Everyone", "loves", "stats.")
c(TRUE, TRUE, FALSE, TRUE)
```

If arguments are vectors, it will be flattened

```{r}
v1 <- c(1,2,3)
v2 <- c(4,5,6)
c(v1,v2)
```

## 6. Computing Basic Statistics

Use one of these functions assuming that x and y are vectors

```{r}
x <- c(1,2,3)
y <- c(6,5,3)
mean(x) # mean
median(x) # median
sd(x) # standard deviation
var(x) # variance
cor(x, y) # correlation
cov(x, y) # covariance
```

They are sensitive to NA, ignore NAs by using `na.rm=TRUE`

```{r}
x <- c(0,1,1,2,3,NA)
mean(x, na.rm=TRUE)

sd(x, na.rm=TRUE)
```

mean and ad understand dataframe and calculate column by column

var function understand dataframe and calculate the covariance matrix between columns

cor and cov of dataframe return correlation and covariance matrix of columns 


## 7. Creating Sequences

Use an `n:m` expression to create the simple sequene n...matrix

```{r}
1:5
```

Use the `seq` function for sequences with an increment other than 1:

```{r}
seq(from=1, to=5, by=2)
```

Use the `rep` funciton to create a series of repeated values

```{r}
rep(1, times=5)
```

Alternatively you can specify a length for the output sequence

```{r}
seq(from=0, to=20, length.out=5)
seq(from=1.0, to= 2.0, length.out=5)
```

## 8. Comparing vectors

Comparison operators: `== != < > <= >=`
Comparing vectors are element by element. Comparing vector and scalar will expand the scalar to a vector.
```{r}
v <- c(3, pi, 4)
w <- c(pi, pi, pi)
v == w
v != w
v < w
v > w
v <= w
v >= w
v == pi
v != pi
```

Use `any` and `all` to summarize

```{r}
v <- c(3,pi,4)
any(v == pi)
any(v != pi)
all(v == pi)
all(v != pi)
all(v >= 2)
```

## 9. Selecting Vector Elemenets

* v[3] for the third element of v1
* Use negative indices to exclude elements
* Use a vector of indices
* Use a logical vector to select elements based on a condition
* Use names to access named elements

**the first element has an index of 1 other than 0**

select multiple elements at once

```{r}
fib <- c(0,1,1,2,3,5,8,13,21,34)
fib[1:3]
fib[4:9]
fib[c(1,2,4,8)]
fib[-1]
fib[-(1:3)]
fib[fib<10]
fib[fib%%2==0]
# select all elemnts greater than median
fib[fib>median(fib)]
# select all elements in the lower and upper 40%
fib[(fib<quantile(fib, 0.2))|(fib>quantile(fib, 0.8))]
# select all elements that exceed +/- 2 standard deviation from the mean
fib[abs(fib-mean(fib))>2*sd(v)]
```

Assign names to vector

```{r}
years <- c(1960, 1964, 1976, 1994)
names(years) <- c("Kennedy", "Johnson", "Carter", "Clinton")
print(years)
years[2]
years["Johnson"]
years[c("Carter", "Johnson")]
```

## 10. Performing Vector Arithmetic

All basic arithmetic operators can be applied to pairs of vectors. 

```{r}
v <- c(11,12,13,14,15)
w <- c(1,2,3,4,5)
v+w
v-w
v*w
v/w
w^v

w+2
w-2
w*2
w/2
2^w
w^2

mean(w)
w-mean(w)

sd(w)
(w-mean(w))/sd(w)

sqrt(w)
log(w)
sin(w)
```

## 11. Getting Operator Precedence Right

|Operator|Meaning|See also|
|:---|:---|:---|
|`[` `[[`|Indexing|Recipe 2.9|
|`::` `:::`|Access variable in a name space||
|TODO|||

## 12. Defining a Function

```{r eval=FALSE}
function(param1, ..., paramn) expression

function(param1, ..., paramn) {
    expr1
    ...
    exprn
}
```

```{r}
cv <- function(x) sd(x)/mean(x)
cv(1:10)
```


